\begin{abstract}
Retinal lesion segmentation is critical for the early diagnosis of Diabetic Retinopathy (DR). However, accurate delineation of minute anomalies—particularly Microaneurysms (MA) and Soft Exudates (SE)—is severely hindered by extreme class imbalance. Furthermore, translating these diagnostic models to real-world rural clinics demands networks capable of operating efficiently on resource-constrained edge devices (e.g., portable retinal scanners) without relying on cloud computation. In this paper, we propose Ghost-CAS-UNet, a highly robust, lightweight deep learning framework designed specifically for edge inference. By leveraging efficient Ghost Modules alongside Coordinate Attention (CA) and constraining input processing to a $256 \times 256$ sliding window, our approach drastically minimizes computational overhead while retaining spatial precision. Evaluated on the widely recognized IDRiD dataset, the proposed architecture leapfrogs heavy baselines, achieving a peak Mean Dice of 74.40\% (a substantial ~23.6\% absolute improvement over standard DeepLabV3+) and demonstrating unprecedented extraction of sparse microaneurysms at 81.42\% Dice. Crucially, this accuracy is coupled with extreme efficiency: the model requires merely 1.98M parameters and an 7.88 MB disk footprint—an approximate $12\times$ reduction compared to ResNet34-backed networks. Utilizing bounded memory ($\sim$684 MB) with 934 ms unoptimized inference latency natively on a Raspberry Pi 4 CPU, our framework represents a pivotal step towards highly accurate, entirely offline CAD screening for point-of-care environments.
\end{abstract}
